### 2022-09-13 main/chapter_searches.tex
add pvalue misinterpretations paragraphs

% p-values
\TODO{no!}~\cite{
schervish1996p,
Cohen1994TheEI,
Amrhein2017TheEI,
goodman2008dirty,
greenland2016no,
Bernstein2016princess,
Trafimow2015ban,
wagenmakers2007practical,
Read2002cls
}
\TODO{examples of use and misuse in hep}

\TODO{p-value footnote:
The chance of a small \pvalue\ is small.
This is true, but not interesting because the value of any \pvalue\ is small.
By construction \pvalues\ are uniformly distributed on the unit interval
under a given hypothesis.
The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
$0.6 < p < 0.65$.
The frequency of $p < \varepsilon$ is identical to the frequency
of $0.6 < p < 0.6 + \varepsilon$.
An appropriate choice of test statistic is therefore necessary for the
\pvalue\ to have meaning.
}

@@

$\cls$~\cite{
zech1988cls,
bock1998lower,
read1997optimal,
Read2002cls,
Murray2010heretic
}
\begin{equation}
\cls
=
\frac{P(s + b)}{P(b)}
=
\frac{
\prob{t \geq \tilde{t}}{s + b}
}{
\prob{t \geq \tilde{t}}{b}
}
.
\end{equation}
These \pvalues\ are sometimes labelled $\clsb = P(s + b)$ and $\clb = P(b)$
Some literature also defines $\clb = 1 - p_b$~\cite{
Cowan:2010js,
pdg2022ynf
},
where $p_b$ is not quite a \pvalue\ if it excludes equality with the test
statistic.

>>>>

Data analysis in modern High Energy Physics does not exactly follow frequentist
ideology, but is a hands-on, practical, and evolving mixture that has adapted
to avoid some the more glaring flaws.
We have seen in Section~\ref{sec:searches_data_analysis} how likelihood ratios
can clearly express our scientific interest in how well competing models
predict data, and how those likelihood can be manipulated in probability
algebra.
We have also seen in Section~\ref{sec:searches_frequentist} how 20th century
writers championed likelihoods not on the data themselves but on inequalities,
forming \pvalues\ and confidence intervals in a reductions that throw away
information and obfuscate the results.
This section reviews how these objects are applied generally in conventional
analysis at the LHC and specifically towards the \atlas\ search that this
thesis presents.

Misinterpretations of \pvalues\ are widespread and well reported~\cite{
schervish1996p,
Cohen1994TheEI,
Amrhein2017TheEI,
goodman2008dirty,
greenland2016no,
Bernstein2016princess,
wagenmakers2007practical,
Trafimow2015ban
}.
For example, the idea of ``probability of compatibility'' with a model~\cite{
HIGG-2018-04
}
appears to be spreading as a good meme~\cite{dawkins1989selfish} in
High Energy Physics~\cite{
HIGG-2018-27,
HIGG-2018-51,
EXOT-2018-08,
HION-2018-19,
HIGG-2017-09,
langford2021combination,
mastrandrea2019searches,
white2019search,
IceCube2013search,
IceCube2014searches,
gerasimov2021new
},
plausibly explained by vague but accurate statements like
``compatible with the SM prediction with a $P$ value of $0.10\%$''~\cite{
lhcb2021test
}
mutating to precise and inaccurate statements like
``a probability of around $0.1\%$ that the data is compatible with the
Standard Model predictions''~\cite{cern2021test}.

There also exists a notion that a small \pvalue\ is significant because
it is a rare fluctuation.
All \pvalues\ are rare ---
by construction they are uniformly distributed, so the
probability of $P(h) < \varepsilon$ is equal to the probability of
$0.4 \leq P(h) < 0.4 + \varepsilon$ or any other $\varepsilon$-sized set
(if the data are sampled under $h$).
All that matters to the interpretation is that an alternative hypothesis makes
the observed result \emph{more probable}.

% CLs
$\cls$~\cite{
zech1988cls,
bock1998lower,
read1997optimal,
Read2002cls,
Murray2010heretic
}
\begin{equation}
\cls
=
\frac{P(s + b)}{P(b)}
=
\frac{
\prob{t \geq \tilde{t}}{s + b}
}{
\prob{t \geq \tilde{t}}{b}
}
.
\end{equation}
These \pvalues\ are sometimes labelled $\clsb = P(s + b)$ and $\clb = P(b)$
Some literature also defines $\clb = 1 - p_b$~\cite{
Cowan:2010js,
pdg2022ynf
},
where $p_b$ is not quite a \pvalue\ if it excludes equality with the test
statistic.

@@

### 2022-09-13 main/draft.tex
add pvalue misinterpretations paragraphs

>>>>

% Read2002cls
% }
\TODO{examples of use and misuse in hep}

\TODO{p-value footnote:
The chance of a small \pvalue\ is small.
This is true, but not interesting because the value of any \pvalue\ is small.
By construction \pvalues\ are uniformly distributed on the unit interval
under a given hypothesis.
The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
$0.6 < p < 0.65$.
The frequency of $p < \varepsilon$ is identical to the frequency
of $0.6 < p < 0.6 + \varepsilon$.
An appropriate choice of test statistic is therefore necessary for the
\pvalue\ to have meaning.
}
